maxent model avec xvalidation (year-based test training)

### 2019-05-01 src/maxent_avec_xvalidation.Rmd
highs and lows..

```{r pre-install & load packages, include=FALSE}
options(java.parameters = "-Xmx1g" )#needs to run before dismo loaded - boosts memory for java to 1g
packages <- c("dismo", "stringr", "rJava", "raster", "ecospat", "bossMaps")
source("../src/ipak.R")
ipak(packages)
```

@@

now to stack average binary maps at depth into one 2D representation...

```{r}

#this below is to get an average score for all layers per year. probably not what you want
ascdepth_list <- list.files(paste0("../output/maxent/predictions/", mthna), pattern = "^binary_mpa_avg_cum_.*asc", full.names = TRUE)

binaryavgstack <- stack(ascdepth_list)

stsum <- sum(binaryavgstack, na.rm = TRUE)
  writeRaster(stsum, (paste0("../output/maxent/predictions/", mthna, "/binary_sum_", mthna, ".asc")), overwrite=TRUE)
  sumplot <- plot(stsum, png(paste0("../output/maxent/predictions/", mthna, "/binary_sum_", mthna, ".png"))) 
  dev.off()
stavg <- calc(binaryavgstack, fun = mean, na.rm = TRUE)
  writeRaster(stavg, (paste0("../output/maxent/predictions/", mthna, "/binary_avg_", mthna, ".asc")), overwrite=TRUE)
  avgplot <- plot(stavg, png(paste0("../output/maxent/predictions/", mthna, "/binary_avg_", mthna, ".png"))) 
  dev.off()
```

@@

```{r}
test <- cellStats(binaryavgstack[[30]], sum)
test
```

>>>>

```{r pre-install & load packages, include=FALSE}
options(java.parameters = "-Xmx1g" )#needs to run before dismo loaded - boosts memory for java to 1g
packages <- c("dismo", "stringr", "rJava", "raster", "ecospat", "bossMaps", "lattice")
source("../src/ipak.R")
ipak(packages)
```

@@

@@

```{r calc no presence cells per layer (avg)} 
ascdepth_list <- list.files(paste0("../output/maxent/predictions/", mthna), pattern = "^binary_mpa_avg_cum_.*asc", full.names = TRUE)
binaryavgstack <- stack(ascdepth_list)
stackstat <- data.frame(matrix(ncol = 2, nrow = nlayers(binaryavgstack)))
stackstat[ ,1] <- c(1:49)
colnames(stackstat) <- c("layer", "cell_sum")

for (lyr in 1:nlayers(binaryavgstack)){
  lyrsum <- cellStats(binaryavgstack[[lyr]], sum)
  stackstat$cell_sum[lyr] <- lyrsum
}

write.csv(stackstat, paste0("../output/maxent/predictions/", mthna, "/cellcount_bydepth_avg.csv"), row.names = FALSE)
```

a crude plot but ok to show slr

```{r graph by depth (avg)}
bydepth <- barchart(layer ~ cell_sum, data = stackstat, ylim = c(49, 1), xlim = c(1, 1600))
trellis.device(device="png",filename = paste0("../output/maxent/predictions/", mthna, "/", mthna, "_depthplot.png"))
print(bydepth)
dev.off()
```

what about the variation from one year to the next

e.g. binary_mpa_st200401041.18.asc

```{r no presence cells per layer per year}
binaryyr_list <- list.files(paste0("../output/maxent/predictions/", mthna), pattern = "^binary_mpa_st.*asc", full.names = TRUE)
yrlayer <- data.frame(matrix(ncol = 2, nrow = length(binaryyr_list)))
colnames(yrlayer) <- c("layer", "year")

for (yrlyr in 1:length(binaryyr_list)){
  yearoflyr <- str_sub(binaryyr_list[yrlyr], start = 47, end = 50) #selects the year from the file name
  yrlayer$year[yrlyr] <- yearoflyr
  yrlayer$layer[yrlyr] <- binaryyr_list[yrlyr]
}

y_list <- split(yrlayer, yrlayer$year)
noyrs <- unique(yrlayer$year)

stackstat_yr <- data.frame(matrix(ncol = length(noyrs) + 1, nrow = nlayers(binaryavgstack)))
stackstat_yr[ ,1] <- c(1:49)
colnames(stackstat_yr) <- c("layer", paste0("y_", noyrs))

for (binyr in (1:length(y_list))){
  lsta <- stack(y_list[[binyr]]$layer)
  yr <- y_list[[binyr]]$year[1]
  for (l in 1:nlayers(lsta)){
    lyrsum <- cellStats(lsta[[l]], sum)
    stcol <- paste0("y_", yr) 
    stackstat_yr[[l, stcol]] <- lyrsum
  }
}

#2005 coming up empty...
